---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
```{r}
#download.file("https://ibm.box.com/shared/static/5wah9atr5o1akuuavl2z9tkjzdinr1lv.csv",              destfile = "recipes.csv", quiet = TRUE)

# download.file("https://ibm.box.com/shared/static/rtqci87zjkgt7jpac9egrln9kifqqxuq.zip",
#               destfile = "/resources/rstudio/recipes-shiny.zip", quiet = TRUE)
# unzip("/resources/rstudio/recipes-shiny.zip")

## Install libraries if not installed
if("rpart" %in% rownames(installed.packages()) == FALSE) {install.packages("rpart", 
    repo = "http://mirror.las.iastate.edu/CRAN/")}
if("rpart.plot" %in% rownames(installed.packages()) == FALSE) {install.packages("rpart.plot", 
    repo = "http://mirror.las.iastate.edu/CRAN/")}

options(scipen = 999) #disable scientific notation
R.version.string
```
# Intro 

This is a sample based on an original version from IBM [Cognitive Builder Faire](https://cognitivebuilder.github.io/) in Seattle on March 26, 2017, by Polong Lin, Joseph Santarcangelo, Ph.D. 

I redid it in RStudio for my own educational purposes.

First, load the data:

```{r}
recipes <- read.csv("recipes.csv")
grep("wasabi", names(recipes), value = TRUE) #Yes
```

Make a simple frequency table to understand a little about the data:

```{r}
t <- base::table(recipes$country) #frequency table
sort(t, decreasing = TRUE)[1:5]
```

Next clean the data

```{r}
recipes$country <- tolower( as.character(recipes$country) )
recipes$country[recipes$country == "china"] <- "chinese"
recipes$country[recipes$country == "france"] <- "french"
recipes$country[recipes$country == "germany"] <- "german"
recipes$country[recipes$country == "india"] <- "indian"
recipes$country[recipes$country == "israel"] <- "jewish"
recipes$country[recipes$country == "italy"] <- "italian"
recipes$country[recipes$country == "japan"] <- "japanese"
recipes$country[recipes$country == "korea"] <- "korean"
recipes$country[recipes$country == "mexico"] <- "mexican"
recipes$country[recipes$country == "scandinavia"] <- "scandinavian"
recipes$country[recipes$country == "thailand"] <- "thai"
recipes$country[recipes$country == "vietnam"] <- "vietnamese"
```

*Problem: Some countries have very few recipes*

Remove data for countries with <50 recipes:

```{r}
t <- sort(base::table(recipes$country), decreasing = T)

filter_list <- names( t[ t >= 50 ] )

before <- nrow(recipes) #number of rows of original df

recipes <- recipes[recipes$country %in% filter_list, ]

after <- nrow(recipes)

print(paste(before - after, "rows removed."))

sort(base::table(recipes$country), decreasing = T)
```

Convert all of the columns into factors (to run the classification model later)

```{r}
recipes[,names(recipes)] <- lapply(recipes[,names(recipes)] , as.factor)
```

## Q: Can we tell that some food is Japanese if it contains rice and soy sauce and wasabi and seaweed?

```{r}
head(recipes[recipes$rice == "Yes" &
        recipes$soy_sauce == "Yes" &
        recipes$wasabi == "Yes" &
        recipes$seaweed == "Yes",1:5])
```

## Q: Which ingredients are most common? Which are the least-used ingredients?¶

Goals:
* count the ingredients across all recipes


```{r}
## Sum the row count when the value of the row in a column is equal to "Yes" (value of 2)
ingred <- unlist(
            lapply( recipes[, names(recipes)], function(x) sum(as.integer(x) == 2))
            )

## Transpose the dataframe so that each row is an ingredient
ingred <- as.data.frame( t( as.data.frame(ingred) ))
                
ing_df <- data.frame("ingredient" = names(ingred), 
                     "count" = as.numeric(ingred[1,])
                    )
                
ing_df_sort <- ing_df[order(ing_df$count, decreasing = TRUE),]
rownames(ing_df_sort) <- 1:nrow(ing_df_sort)
head(ing_df_sort)
```

difference among countries

```{r}
by_country_norm <- aggregate(recipes, 
                        by = list(recipes$country), 
                        FUN = function(x) round(sum(as.integer(x) == 2)/
                                                length(as.integer(x)),4))
#Remove the unnecessary column "country"
by_country_norm <- by_country_norm[,-2]

#Rename the first column into "country"
names(by_country_norm)[1] <- "country"
                                              
head(by_country_norm[,1:5])
```

# Find specific ethnic foods

```{r}
region <- "canada" #try: italian

regiondata <- by_country_norm[by_country_norm$country == region,]
head(t(sort(regiondata, decreasing = TRUE)))
```

top 5 most popular ingredients in each country

```{r}
for(nation in by_country_norm$country[1:7]){
    x <- sort(by_country_norm[by_country_norm$country == nation,], decreasing = TRUE)
    cat(c(toupper(nation)))
    cat("\n")
    cat(paste0(names(x)[2:6], " (", round(x[2:6]*100,0), "%) "))
    cat("\n")
    cat("\n")
}
```

k-means clustering
```{r}
n_cluster = 5
df_cluster <- by_country_norm
k <- kmeans(df_cluster[,-1], n_cluster)
df_cluster$cluster <- k$cluster
for(i in seq( 1, n_cluster )){
    i <- as.character(i)
    cat(paste0("[Cluster ", i, "]----------------------------------------------------------"))
    cat("\n")
        
    print(paste0(as.character(df_cluster[df_cluster$cluster == i,]$country)))
    cat("\n")
    }



```

# Decision Trees

```{r}
# Run this

## Install libraries if not installed
if("rpart" %in% rownames(installed.packages()) == FALSE) {install.packages("rpart", 
    repo = "http://mirror.las.iastate.edu/CRAN/")}
if("rpart.plot" %in% rownames(installed.packages()) == FALSE) {install.packages("rpart.plot", 
    repo = "http://mirror.las.iastate.edu/CRAN/")}


## Load libraries
library(rpart)
library(rpart.plot)
```


```{r}
#Choose one:
head(recipes[,1:5])
#str(recipes)
#table(recipes$country)
```


But American data dominiates so much, that we can't do good decision trees.

```{r}
## Select subset of countries
bamboo_tree <- rpart(formula = country ~ ., 
                     data = recipes[recipes$country %in% c("korean", 
                                                           "japanese", 
                                                           "chinese", 
                                                           "thai",
                                                           "indian"),], 
                     method ="class")
```


```{r}
# Run this

## Plot the East Asian + Indian model
## run "?rpart.plot" if you want to see the arguments for rpart.plot
rpart.plot(bamboo_tree, type = 3, fallen.leaves = F, extra = 2, under = TRUE, cex = 0.75, varlen = 0, faclen = 0)

```

more summaries

```{r}
names(bamboo_tree)
```



more

```{r}
bamboo <- recipes[recipes$country %in% c("korean", "japanese", "chinese",
                                              "thai", "indian"),]
table(as.factor(as.character(bamboo$country)))
```

```{r}
sample_n <- 100
## Take n recipes from each country
set.seed(4) #Set random seed
korean <- bamboo[sample(which(bamboo$country == "korean") , sample_n), ]
japanese <- bamboo[sample(which(bamboo$country == "japanese") , sample_n), ]
chinese <- bamboo[sample(which(bamboo$country == "chinese") , sample_n), ]
thai <- bamboo[sample(which(bamboo$country == "thai") , sample_n), ]
indian <- bamboo[sample(which(bamboo$country == "indian") , sample_n), ]

#Create the dataframe
bamboo_test <- rbind(korean,japanese, chinese, thai, indian)
bamboo_train <- bamboo[!(rownames(bamboo) %in% rownames(bamboo_test)),]

## Check that we have 100 recipes from each cuisine
table(as.factor(as.character(bamboo_test$country)))

## Check that we have subtracted 100 recipes from each cuisine
table(as.factor(as.character(bamboo_train$country)))
```


```{r}
#Train on the bamboo_train data
bamboo_tree_pred <- rpart(formula = country ~ ., 
                     data = bamboo_train[bamboo_train$country %in% c("korean", 
                                                           "japanese", 
                                                           "chinese", 
                                                           "thai",
                                                           "indian"),], 
                     method ="class")

```

plot it

```{r}
#Plot the trained tree
rpart.plot(bamboo_tree_pred, type = 3, extra = 2, fallen.leaves = F, under = TRUE, 
           cex = 0.75, varlen = 0, faclen = 0)
```



# Create a confusion matrix:
compare the predictions (from the model) with the actual cuisine labels (from the recipes)

```{r}
bamboo_fit <- predict(bamboo_tree_pred, subset(bamboo_test, select=-c(country)), type = "class")
bamboo_tab <- table(paste(as.character(bamboo_fit),"_pred", sep =""), 
                    paste(as.character(bamboo_test$country),"_true", sep =""))
bamboo_tab
```


# Intrepreting the table:
* each column is the actual recipe's label (from the website)
* each value represents the percentage of country X's recipes correctly labeled as country X
* Example: 63.3% of the Chinese recipes were correctly predicted as "Chinese", incorrectly predicted 30.0% of the time as "Korean", and incorrectly predicted 6.7% as "Japanese"

# Discussion:

* Which cuisines were likely to be mis-classified or correctly classified?
* Which cuisines seem to be mistaken for which other cuisines?


# Deployment

```{r}
#Creates an empty recipe, "my_recipe"
my_recipe <- recipes[1,]
my_recipe[my_recipe[1,]== "Yes"] <- "No"
my_recipe$country <- NULL

#Function to find an ingredient 
ing_find <- function(ing){
    if(length(grep(tolower(ing),names(recipes), value = TRUE)) > 0){
        grep(tolower(ing),names(recipes), value = TRUE)
        }
    else{paste(ing,"was not found.")}
    }

#Function to predict using the bamboo_tree_pred model, by default
recipe_predict <- function(inputrecipe, model = bamboo_tree_pred){
    p <- as.character(predict(model, inputrecipe, type = "class"))
    paste("Hmm... your recipe tastes rather...", toupper(p))
}
```



# Check your ingredient names:

First, check that your ingredient exists in the list of ingredients:

```{r}
ing_find("wasabi")
ing_find("tea")
```


add ingredients
```{r}
#Creates an empty recipe, "my_recipe"
my_recipe <- recipes[1,]
my_recipe[my_recipe[1,]== "Yes"] <- "No"
my_recipe$country <- NULL

### YOUR CODE BELOW ###
my_recipe$rice <- "Yes"
my_recipe$wasabi <- "Yes"
my_recipe$soy_sauce <- "Yes"

recipe_predict(my_recipe, model = bamboo_tree) #or olive_tree or baobab_tree
```


bit.ly/recipes-shiny


# references and notes


Recipes dataset originally from:
Ahn, Yong-Yeol, et al. "Flavor network and the principles of food pairing." Scientific reports 1 (2011).
Dataset on ingredient-compounds also available here:
recipes.csv (64.2 MB): http://bit.ly/recipesdata


